---
title: "Documentation about the package"
author: "Ana Ezquerro"
date:  "Versión `r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: yes
    latex_engine: lualatex
    fig_caption: yes
    toc: yes
    highlight: tango
    df_print: kable
    citation_package: biblatex
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    number_sections: true
    theme: paper
    highlight: tango
  prettydoc::html_pretty:
    toc: true
    df_print: paged
    number_sections: true
    theme: hpstr
    highlight: github
  md_document:
    variant: markdown_github
fontsize: 12pt
geometry: margin=0.7in
classoption: a4paper
documentclass: article
header-includes:
- \usepackage{sfmath}
- \renewcommand*\familydefault{\sfdefault}
- \renewcommand{\baselinestretch}{1.2}
- \setlength{\parskip}{1em}
- \usepackage{xcolor}
- \input{confi.tex}
always_allow_html: yes
bibliography: references.bib
biblio-style: bwl-FU
linkcitations: true
linkcolor: blue
lang: es
---

```{r, echo=F, warning=F, message=F}
# Cargamos las funciones
knitr::opts_chunk$set(fig.align='center', fig.width = 10, 
                      fig.height = 8, message=F, comment='', warning=F)
eval(parse("plot-tools.R", encoding="UTF-8"))
eval(parse("auto-fit.R", encoding="UTF-8"))
eval(parse("auto-select.R", encoding="UTF-8"))
eval(parse("forecasting.R", encoding="UTF-8"))

# Time series analysis
library(fpp2)
library(tseries)
library(TSA)
library(seastests)
library(forecast)

# Plot libraries
library(plotly)

# Auxiliars
library(prettydoc)
library(stringi)
library(stringr)
library(polynom)
library(parallel)

display <- function(fig, name, width=800, height=400) {
  
  if (is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))) {
    return(fig)
  }
  if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "latex") {
    figpath <- paste0('figures/', name, ".png")
    save_image(fig, figpath, width=width, height=height)
    return(knitr::include_graphics(figpath))
  }
  if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "html") {
    fig <- fig %>% layout(width=840, height=700)
    return(fig)
  }
}

dir.create("figures", showWarnings=F)
```



\newpage

# Function for automatic fitting ARIMA or ARIMAX models (`auto.fit.arima`)


**Description**: Obtains a valid fitted ARIMA or ARIMAX model that optimizes a selected information criterion and satisfies that:

1) All coefficients are statistically significtive.
2) Model residuals have zero mean and are independent.

**Returns**:

a. Fitted ARIMA or ARIMAX model (`Arima` object) if it exists and it can be optimized.
b. `NA` in case it does not exists and/or cannnot be optimized.

```{r, eval=F}
auto.fit.arima(serie, xregs = NULL,  ic = c("aicc", "aic", "bic"), d = NA, 
               D = NA, alpha = 0.05, show_info = TRUE, plot_result = FALSE)
```

**Arguments**:

- `serie` [`ts`]: Univariate time series to fit an ARIMA model with. If `xregs` is not `NULL` it will act as the dependent variable.
- `xregs` [`mts`]: Matrix of time series that act as covariates in an ARIMAX model. By default, `NULL`, meaning that an ARIMA will be fitted with `serie`.
- `ic` [`character`]: Information criterion to be used in ARIMA orders' selection. Options available are the same as in `forecast::auto.arima()`  function: `ic`, `bic` or `aicc`.
- `d` [`numeric`]: Value of the diferentiation order. By default `NA`, thus there is no restriction about the $d$ order value.
- `D` [`numeric`]: Value fo the seasonal diferentiation order. By default `NA`, thus there is no restriction about the $D$ order value.
- `alpha` [`numeric`]: Significance level of hypothesis tests used for checking independence, zero mean and normality of residuals, and significance  of estimated coefficients.
- `show_info` [`boolean`]: Displaying or not displaying the historical of fitted models.
- `plot_result` [`boolean`]: Returning or not with the fitted model a sequential plot of time series and residuals.

**Observations**:

- To check residuals independence the Ljung-Box (`Box.test()`) is used. The number of lags is selected from the length of the time serie and testing the stationarity of it (see function `ljunbox_lag()`).
- To check null mean of residuals the `t.test()` is used.
- To check normality of residuals we use the Jarque-Bera (`jarque.bera.test()`) and the Shapiro-Wilks tests (`shapiro.test()`).
- Considered models will always have a regular differentiation order less or equal to 3 ($d \leq 3$) and a seasonal differentiation order less or equal to 2 ($D \leq 2$).

**Example**: Influenza evolution in Catalonia. 

```{r}
dat <- read.csv("data/evolucion_gripe_covid.csv")               # read data
sdgripal <- ts(dat$sdgripal, start=c(2020, 40), frequency=52)   # convert to ts
model_gripal <- auto.fit.arima(sdgripal, plot_result = TRUE)
```


```{r}
# display series sequential plot
display(model_gripal$fig_serie, "serie gripe", width=1000, height=800)
```


```{r}
# display residuals sequential plot
display(model_gripal$fig_residuals, "residuals gripe", width=1000, height=800)
```

**Example**: [Monthly level of Co2 measured in the Observatory of Manua Loa (Hawaii)](ftp://ftp.cmdl.noaa.gov/ccg/co2/trends/co2_mm_mlo.txt). Time series starts in March 1958.

```{r}
# read and parse data
co2 <- ts(scan('data/co2MaunaLoa.dat'), start=c(1958, 3), frequency=12)

# apply the automatic fitting
model_co2 <- auto.fit.arima(co2, ic="aicc", plot_result=TRUE)
```


```{r}
# display sequential plot of the time series
display(model_co2$fig_serie, "serie co2", width=1000, height=800)
```


```{r}
# display sequential plot of the residuals
display(model_co2$fig_residuals, "residuals co2", width=1000, height=800)
```

\newpage

# Function for automatic covariates selection and lags in Dynamic Regression models (`drm.select`)


**Description**: Implementation of our new approach of covariates selection in dynamic regression models where each covariate  is introduced lagged $k\geq 0$ moments. For more information about this method, check the paper attached with this file.

**Returns**:

- [`Arima`] Fitted ARIMAX model with significant covariates selected.
- [`Arima`] Fitted ARIMA model with the dependent variable in case no covariate is selected.

```{r, eval=F}
drm.select(serie, xregs, ic=c('aic', 'bic', 'aicc', alpha=0.05, 
                              st_method='auto.arima', show_info=T, ndiff)
```

**Arguments**:



 - `serie` [`ts`]: Univariate time series which acts as the dependent variable.
 - `xregs` [`mts`]: Set of covariate candidates to model the behavior of `serie`.
 - `ic` [`character`]: Information criterion used for covariate selection and model comparisons. The possibilities are: `aic`, `bic` and `aicc`. By default, th AICc is used.
 - `alpha` [`numeric`]: Signficance level for hypothesis tests of stationary and coefficients significance. By default it is set to $0.05$.
 - `st_method` [`character`]: Method used for checking stationary of a time series. If it is `auto.arima`, the function `forecast::auto.arima()` is used and the differentiation order is checked. If it is `adf.test`, the Dickey-Fuller test is used.
 - `show_info` [`boolean`]: Displaying or not displaying the historical of added covariates.
 - `ndiff` [`numeric`]: Internal argument (**do not use**) to apply regular differentiations to all data when no ARIMAX model can be fitted with stationary errors.


*Note*: Information about inidividual model fitting of each covariate will not be displayed. 

**Example**: Logarithm of weekly sales and *Bluebird* chips price in New Zeland. The observation period is 104 weeks long (from 20/09/1988 to 10/09/2000). 

```{r}
load("data/patatas.dat")                    # read data
sales <- patatas[,1]
xregs <- ts(matrix(patatas[,2]))
colnames(xregs) <- c('price')

chips_model <- drm.select(sales, xregs)     # apply selection method
```

**Example**: Time series about Microsoft *stock*. 

```{r}
microsoft <- read.csv('data/microsoft-stock.csv')     # read data
close_price <- ts(microsoft$Close)                    # dependent variable
xregs <- ts(microsoft[, c('Open', 'High', 'Low', 'Volume')])

# fit model
fitted_model <- drm.select(close_price, xregs)
```




**Example**:  Exitus cases in Spain due to COVID19 virus, considering the following covariates:

-   Confirmed cases and recovered cases in Spain. 
-   Confirmed, exitus and recovered cases in France.
-   Confirmed, exitus and recovered cases in Portugal.
-   Confirmed, exitus and recovered cases in England.


```{r}
# read data
confirmed <- read.csv("data/covid-global-confirmed-bycountry.csv")
deaths <- read.csv("data/covid-global-deaths-bycountry.csv")
recovered <- read.csv("data/covid-global-recovered-bycountry.csv")

# Spain data
deaths_spain <- ts(deaths$Spain, frequency=7)
confirmed_spain <- ts(confirmed$Spain, frequency=7)
recovered_spain <- ts(recovered$Spain, frequency=7)

# France data
deaths_france <- ts(deaths$France, frequency=7)
confirmed_france <- ts(confirmed$France, frequency=7)
recovered_france <- ts(recovered$France, frequency=7)

# England data
deaths_england <- ts(deaths$United.Kingdom, frequency=7)
confirmed_england <- ts(confirmed$United.Kingdom, frequency=7)
recovered_england <- ts(recovered$United.Kingdom, frequency=7)

# Portugal data
deaths_portugal <- ts(deaths$Portugal, frequency=7)
confirmed_portugal <- ts(confirmed$Portugal, frequency=7)
recovered_portugal <- ts(recovered$Portugal, frequency=7)

xregs <- ts(cbind(confirmed_spain, recovered_spain, 
                  deaths_france, confirmed_france, recovered_france,
                  deaths_england, confirmed_england, recovered_england,
                  deaths_portugal, confirmed_portugal, recovered_portugal),
            frequency=7)

# fit model
fitted_model <- drm.select(deaths_spain, xregs)
```


\newpage

# Point predictions at horizon $h$ and prediction intervals (`forecast_model`)

Once a dynamic regression model has been fitted, we can make point predictions at horizon $h$ with their respective prediction intervals given the `Arima` object from `drm.select()`.

In order to make point predictions at horizon $h$ from a DRM it is necessary to have covariates' predictions at the same horizon $h$. These value can be obtained from:

- If the lag $r$ of the covariate is grater or equal $h$, the $remaining$ values of the original time serie.
- If the lag $r$ of the covariate is less than $h$, by fitting an ARIMA model with the covariate observations and predicting $h-r$ values.



**Description**: Given an horizon $h$, compute point predictions and prediction intervals based in a DRM computed by the function `drm.select()`.


**Returns**: Point predictions and prediction intervals in original scale.

```{r, eval=F}
forecast_model(serie, xregs, fitted_model, h, mode='bootstrap', levels=c(80, 90))
```

**Argumentos**:

- `serie` [`ts`]: Univariate time series which acts as the dependent variable. It must be the same variable used in `drm.select()`.
- `xregs` [`mts`]: Matrix time series with the selected covariates. It must be the same matrix used in `drm.select()`.
- `fitted_model` [`Arima`]: Fitted ARIMAX model obtained.
- `h` [`numeric`]: Horizon value for forecasting. 
- `mode` [`character`]: Type of prediction intervals.: based on residuals normality (`norm`) or via bootstrap (`bootstrap`). By default `bootstrap` .
- `levels` [`vector`]: Numerical vector of the confidence intervals.

**Example**: 

```{r}
# read data and fit model
load("data/patatas.dat")
sales <- patatas[,1]   
xregs <- ts(matrix(patatas[,2]))
colnames(xregs) <- c('price')
chips_model <- drm.select(sales, xregs, show_info=F)

# point predictons
preds <- forecast_model(sales, xregs, chips_model, h=10, mode='bootstrap')
display(plot_forecast(preds), name='preds_chips')
```



