---
title: "Documentation and examples of use of the `dynamic-arimax` repository"
author: "Ana Ezquerro"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    number_sections: true
    theme: paper
    highlight: tango
  pdf_document:
    number_sections: yes
    latex_engine: lualatex
    fig_caption: yes
    toc: yes
    highlight: tango
    df_print: kable
    citation_package: biblatex
  prettydoc::html_pretty:
    toc: true
    df_print: paged
    number_sections: true
    theme: hpstr
    highlight: github
fontsize: 12pt
geometry: margin=0.7in
classoption: a4paper
documentclass: article
header-includes:
- \usepackage{sfmath}
- \renewcommand*\familydefault{\sfdefault}
- \renewcommand{\baselinestretch}{1.2}
- \setlength{\parskip}{1em}
- \usepackage{fancyvrb}
- \usepackage{xcolor}
- \input{confi.tex}
always_allow_html: yes
biblio-style: bwl-FU
linkcitations: true
linkcolor: blue
---


```{r, echo=F, warning=F, message=F}
knitr::opts_chunk$set(fig.align='center', fig.width = 10, 
                      fig.height = 8, message=F, comment='', warning=F)
eval(parse("../plot-tools.R", encoding="UTF-8"))
eval(parse("../auto-fit.R", encoding="UTF-8"))
eval(parse("../auto-select.R", encoding="UTF-8"))
eval(parse("../forecasting.R", encoding="UTF-8"))
eval(parse('../arima-simulation.R', encoding='UTF-8'))

library(fpp2)
library(tseries)
library(TSA)
library(seastests)
library(forecast)

library(plotly)
library(forecast)

library(prettydoc)
library(stringi)
library(stringr)
library(polynom)
library(parallel)

display <- function(fig, name, width=800, height=400) {
  if (is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))) {
    return(fig)
  }
  if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "latex") {
    figpath <- paste0('figures/', name, ".pdf")
    save_image(fig, figpath, width=width, height=height)
    return(knitr::include_graphics(figpath))
  }
  if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "html") {
    fig <- fig %>% layout(width=840, height=700)
    return(fig)
  }
}

dir.create("figures", showWarnings=F)
```

# Table of Contents


1. [`auto.fit.arima()`](#auto.fit.arima)
2. [`drm.select()`](#drmselect)
3. [`forecast_model()`](#forecast)


# `auto.fit.arima()`:  Function for automatic fitting ARIMA or ARIMAX models {#autofitarima}



**Description**: Obtains a valid fitted ARIMA or ARIMAX model that
optimizes a selected information criterion and satisfies that:

1.  All coefficients are statistically significative.
2.  Model residuals have zero mean and are independent.

**Returns**:

1.  Fitted ARIMA or ARIMAX model (`Arima` object) if it exists and it
    can be optimized.
2.  `NA` in case it does not exists and/or can not be optimized.

```{r, eval=F}
auto.fit.arima(serie, xregs = NULL,  ic = c("aicc", "aic", "bic"), d = NA, 
               D = NA, alpha = 0.05, show_info = TRUE, plot_result = FALSE)
```

**Arguments**:

-   `serie` [`ts`]: Univariate time series to fit an ARIMA model with.
    If `xregs` is not `NULL` it will act as the dependent variable.
-   `xregs` [`mts`]: Matrix of time series that act as covariates in
    an ARIMAX model. By default, `NULL`, meaning that an ARIMA will be
    fitted with `serie`.
-   `ic` [`character`]: Information criterion to be used in ARIMA
    orders’ selection. Options available are the same as in
    `forecast::auto.arima()` function: `ic`, `bic` or `aicc`.
-   `d` [`numeric`]: Value of the diferentiation order. By default
    `NA`, thus there is no restriction about the $d$ order value.
-   `D` [`numeric`]: Value fo the seasonal diferentiation order. By
    default `NA`, thus there is no restriction about the $D$ order
    value.
-   `alpha` [`numeric`]: Significance level of hypothesis tests used
    for checking independence, zero mean and normality of residuals, and
    significance of estimated coefficients.
-   `show_info` [`boolean`]: Displaying or not displaying the
    historical of fitted models.
-   `plot_result` [`boolean`]: Returning or not with the fitted model
    a sequential plot of time series and residuals.

**Observations**:

-   To check residuals independence the Ljung-Box (`Box.test()`) is
    used. The number of lags is selected from the length of the time
    serie and testing the stationarity of it (see function
    `ljunbox_lag()`).
-   To check null mean of residuals the `t.test()` is used.
-   To check normality of residuals we use the Jarque-Bera
    (`jarque.bera.test()`) and the Shapiro-Wilks tests
    (`shapiro.test()`).
-   Considered models will always have a regular differentiation order
    less or equal to 3 ($d\leq 3$) and a seasonal differentiation order
    less or equal to 2 ($D\leq 2$).

**Example**: Influenza evolution in Catalonia.
```{r}
dat <- read.csv("../data/evolucion_gripe_covid.csv")               # read data
sdgripal <- ts(dat$sdgripal, start=c(2020, 40), frequency=52)   # convert to ts
model_gripal <- auto.fit.arima(sdgripal, plot_result = TRUE)
```

```{r}
# display series sequential plot
display(model_gripal$fig_serie, "serie gripe", width=1000, height=800)
```


```{r}
# display residuals sequential plot
display(model_gripal$fig_residuals, "residuals gripe", width=1000, height=800)
```


**Example**: [Monthly level of Co2 measured in the Observatory of Manua
Loa (Hawaii)](ftp://ftp.cmdl.noaa.gov/ccg/co2/trends/co2_mm_mlo.txt).
Time series starts in March 1958.

```{r}
# read and parse data
co2 <- ts(scan('../data/co2MaunaLoa.dat'), start=c(1958, 3), frequency=12)

# apply the automatic fitting
model_co2 <- auto.fit.arima(co2, ic="aicc", plot_result=TRUE)
```


```{r}
# display sequential plot of the time series
display(model_co2$fig_serie, "serie co2", width=1000, height=800)
```

```{r}
# display sequential plot of the residuals
display(model_co2$fig_residuals, "residuals co2", width=1000, height=800)
```


# `drm.select()`: Function for automatic covariates selection and lags in Dynamic Regression models {#drmselect}

**Description**: Implementation of our new approach of covariates
selection in dynamic regression models where each covariate is
introduced lagged $k\geq 0$ moments. For more information about this
method, check the paper attached with this file.

**Returns**:

a. `Arima`: Fitted ARIMAX model with significant covariates selected.
b. `Arima`: Fitted ARIMA model with the dependent variable in case no covariate is selected.

```{r, eval=F}
drm.select(serie, xregs, ic=c('aic', 'bic', 'aicc', alpha=0.05, 
                              st_method='auto.arima', show_info=T, ndiff)
```

**Arguments**:

-   `serie` [`ts`]: Univariate time series which acts as the dependent
    variable.
-   `xregs` [`mts`]: Set of covariate candidates to model the behavior
    of `serie`.
-   `ic` [`character`]: Information criterion used for covariate
    selection and model comparisons. The possibilities are: `aic`, `bic`
    and `aicc`. By default, th AICc is used.
-   `alpha` [`numeric`]: Significance level for hypothesis tests of
    stationary and coefficients significance. By default it is set to
    0.05.
-   `st_method` [`character`]: Method used for checking stationary of
    a time series. If it is `auto.arima`, the function
    `forecast::auto.arima()` is used and the differentiation order is
    checked. If it is `adf.test`, the Dickey-Fuller test is used.
-   `show_info` [`boolean`]: Displaying or not displaying the
    historical of added covariates.
-   `ndiff` [`numeric`]: Internal argument (**do not use**) to apply
    regular differentiations to all data when no ARIMAX model can be
    fitted with stationary errors.

*Note*: Information about inidividual model fitting of each covariate
will not be displayed.

**Example**: Logarithm of weekly sales and *Bluebird* chips price in New
Zeland. The observation period is 104 weeks long (from 20/09/1988 to
10/09/2000).

```{r}
load("../data/patatas.dat")
sales <- patatas[,1]   
xregs <- ts(matrix(patatas[,2]))
colnames(xregs) <- c('price')
chips_model <- drm.select(sales, xregs)
```


**Example**: Time series about Microsoft *stock*.

```{r}
microsoft <- read.csv('../data/microsoft-stock.csv')
close_price <- ts(microsoft$Close)   # variable respuesta
xregs <- ts(microsoft[, c('Open', 'High', 'Low', 'Volume')])
fitted_model <- drm.select(close_price, xregs)
```


**Example**:  Exitus cases in Spain due to COVID19 virus, considering the following covariates:

-   Confirmed cases and recovered cases in Spain. 
-   Confirmed, exitus and recovered cases in France.
-   Confirmed, exitus and recovered cases in Portugal.
-   Confirmed, exitus and recovered cases in England.

```{r}
# read data
confirmed <- read.csv("../data/covid-global-confirmed-bycountry.csv")
deaths <- read.csv("../data/covid-global-deaths-bycountry.csv")
recovered <- read.csv("../data/covid-global-recovered-bycountry.csv")

# Spain data
deaths_spain <- ts(deaths$Spain, frequency=7)
confirmed_spain <- ts(confirmed$Spain, frequency=7)
recovered_spain <- ts(recovered$Spain, frequency=7)

# France data
deaths_france <- ts(deaths$France, frequency=7)
confirmed_france <- ts(confirmed$France, frequency=7)
recovered_france <- ts(recovered$France, frequency=7)

# England data
deaths_england <- ts(deaths$United.Kingdom, frequency=7)
confirmed_england <- ts(confirmed$United.Kingdom, frequency=7)
recovered_england <- ts(recovered$United.Kingdom, frequency=7)

# Portugal data
deaths_portugal <- ts(deaths$Portugal, frequency=7)
confirmed_portugal <- ts(confirmed$Portugal, frequency=7)
recovered_portugal <- ts(recovered$Portugal, frequency=7)

xregs <- ts(cbind(confirmed_spain, recovered_spain, 
                  deaths_france, confirmed_france, recovered_france,
                  deaths_england, confirmed_england, recovered_england,
                  deaths_portugal, confirmed_portugal, recovered_portugal),
            frequency=7)

# fit model
fitted_model <- drm.select(deaths_spain, xregs)
```


Note here that that the selected covariates are: `confirmed_spain`, `recovered_portugal`, `recovered_france`, `confirmed_france`, `confirmed_portugal` and `recovered_spain`. Othe covariates appear in the output of the function, but their coefficients are set to zero (since the function `auto.fit.arima` checked that they were not significative), thus, they are not really selected.

# `forecast_model()`: Point predictions at horizon $h$ and prediction intervals {#forecast}


Once a dynamic regression model has been fitted, we can make point
predictions at horizon $h$ with their respective prediction intervals
given the `Arima` object from [`drm.select()`](#drmselect).

In order to make point predictions at horizon $h$ from a DRM it is
necessary to have covariate predictions at the same horizon $h$. These
value can be obtained from:

-   If the lag $r$ of the covariate is grater or equal $h$, the
    *remaining* values of the original time serie.
-   If the lag $r$ of the covariate is less than $h$, by fitting an
    ARIMA model with the covariate observations and predicting $h-r$
    values.

**Description**: Given an horizon $h$, compute point predictions and
prediction intervals based in a DRM computed by the function
[`drm.select()`](#drmselect).

**Returns**: Point predictions and prediction intervals in original
scale.

```{r, eval=F}
forecast_model(serie, xregs, fitted_model, h, mode='bootstrap', levels=c(80, 90))
```

**Argumentos**:

-   `serie` [`ts`]: Univariate time series which acts as the dependent
    variable. It must be the same variable used in [`drm.select()`](#drmselect).
-   `xregs` [`mts`]: Matrix time series with the selected covariates.
    It must be the same matrix used in [`drm.select()`](#drmselect).
-   `fitted_model` [`Arima`]: Fitted ARIMAX model obtained.
-   `h` [`numeric`]: Horizon value for forecasting.
-   `mode` [`character`]: Type of prediction intervals.: based on
    residuals normality (`norm`) or via bootstrap (`bootstrap`). By
    default `bootstrap` .
-   `levels` [`vector`]: Numerical vector of the confidence intervals.

**Example**:

```{r}
# read data and fit model
load("../data/patatas.dat")
sales <- patatas[,1]   
xregs <- ts(matrix(patatas[,2]))
colnames(xregs) <- c('price')
chips_model <- drm.select(sales, xregs, show_info=F)

# point predictions
preds <- forecast_model(sales, xregs, chips_model, h=10, mode='bootstrap')
display(plot_forecast(preds), name='preds_chips')
```
